Individual and Differential Harm in Redistricting

### 2025-08-25 paper/harm.qmd
cuts

To allow for this broader scope, we replace the fixed election $v$, counterfactual districting plan $\tilde q$, and individual voter $i$ with random variables $V\in\Elecs$, $\tilde Q\in\Plans$, and $I\in\Voters$.
The random variable $V$ represents a statistical model for possible future elections---one realization could correspond to a heavily Democratic-favoring midterm year, while another could involve incumbents under-performing particularly badly.
The random variable $\tilde Q$ represents a model for potential counterfactual districting plans.
Rarely would an analyst have a single counterfactual plan in mind: different realizations of $\tilde Q$ correspond to different districting plans that are permissible under the relevant districting criteria.
Finally, $I$ is defined to be a voter taken uniformly at random from $\Voters$, with $I$ independent of $V,\tilde Q$.

@@

Single-member districts are often motivated by the value in representation to specific communities with shared interests.
Preserving "communities of interest" is one of the most common statutory and constitutional requirements imposed by states on the plan-drawing process [@ncslcriteria].
It is therefore natural to consider not just the average harm to voters done by a plan, but also how this harm is distributed across certain groups of voters, defined by race, party, or other attributes.

>>>>

To allow for this broader scope, we replace the fixed election $v$, counterfactual districting plan $\tilde q$, and individual voter $i$ with random variables $V\in\Elecs$, $\tilde Q\in\Plans$, and $I\in\Voters$.
The random variable $V$ represents a statistical model for possible future elections---one realization could correspond to a heavily Democratic-favoring midterm year, while another could involve incumbents under-performing particularly badly.
The random variable $\tilde Q$ represents a model for potential counterfactual districting plans.
Rarely would an analyst have a single counterfactual plan in mind: different realizations of $\tilde Q$ could correspond to different plans that respect the relevant districting criteria.
Finally, $I$ is defined to be a voter taken uniformly at random from $\Voters$, independently of $V,\tilde Q$.

@@

Single-member districts are often motivated by the value in representation to specific communities with shared interests.
<!-- Preserving "communities of interest" is one of the most common statutory and constitutional requirements imposed by states on the plan-drawing process [@ncslcriteria]. -->
It is therefore natural to consider not just the average harm to voters done by a plan, but also how this harm is distributed across certain groups of voters, defined by race, party, or other attributes.
